---
title: "CASA005assign"
author: "YIRAN"
date: "2021/1/8"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#CASA005

-------
## 1. introduction
This is a instruction for Using Uber data to visualize spatial inequality in London.The main work is:
1. Bring in Uber data and project it to London borough.
2. Combine the economic and social conditions of London borough with London borough layer to Visualize economic differences
3. Make interactive maps

## 2. import datasets and shapefiles
### 2.1 import packages
Import most of the packages which will be used.
```{r }
library(knitr)
library(ggplot2)
library(sf)
library(tmap)
library(tmaptools)
library(tidyr)
library(tidyverse)
library(here)
library(osmar)
library(osmdata)
library(maptools)
library(xml2)
library(RColorBrewer)
```
### 2.2 import data and data processing
#### 2.2.1 Uber data
First, import the Q1 Uber data.Got the dataset from [Uber movement](https://movement.uber.com/explore/london/speeds/query?dt[tpb]=ALL_DAY&dt[wd;]=1,2,3,4,5,6,7&dt[dr][sd]=2020-03-01&dt[dr][ed]=2020-03-31&ff=&lang=en-GB), choose the city London then the porduct type is speed. Data-Time range is 1/3/2020 - 31/3/2020. And I have put the dataset in my [github](https://github.com/Yiran-Dou/CASA005FINAL).
```{r}
Uber <- read_csv("movement-speeds-quarterly-by-hod-london-2020-Q1.csv",
                 locale = locale(encoding = "latin1"),
                 na = "n/a")
```
check the type of data:
```{r}
class(Uber)
Datatypelist <- Uber%>% 
  summarise_all(class)%>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

```
In order to roughly observe Uber data, plot the average speed change at each time of the day.
```{r}
mean(Uber$speed_mph_mean)
median(Uber$speed_mph_mean)   ##mean and median of all data

Ubermean <-aggregate(.~hour_of_day,data=Uber,mean)   ## mean speed of each hour of day
```


```{r pressure, echo=FALSE}
library(png)
sp<- plot(Ubermean$speed_mph_mean~Ubermean$hour_of_day,
          type='l',
          xlab='time of day',xaxt = "n",
          ylab='speed mph mean',col='blue')

abline(h=mean(Uber$speed_mph_mean),col='red')   ## add the mean speed line
grid(6,5, lwd = 1,col = "black")
axis(1,0:23,0:23)  
legend("topright",                                   
      legend=c("speed change","mean speed"),        
      col=c("blue","red"),                 
      lty=1,lwd=1) 
```
As the Uber data used the OSM node ID, we need to transfer it to the coords data for Correctly mark those on the map.
ways
```{r}
id <- Uber$osm_way_id
ways <- opq_osm_id(type ='way', id = id)%>%
  opq_string()%>%
  osmdata_sp()
```
start
```{r}
start <- Uber$osm_start_node_id
start <- opq_osm_id(type ='node', id = start)%>%
  opq_string()%>%
  osmdata_sp()
```
end
```{r}
end <- Uber$osm_end_node_id
end <- opq_osm_id(type ='node', id = end)%>%
  opq_string()%>%
  osmdata_sp()
```
Convert coordinate system:
```{r}

startpoint<- as.data.frame(start[["osm_points"]]@coords)
class(startpoint)
stp<- startpoint%>%
  st_as_sf(., coords = c("lon", "lat"), 
           crs = 4326) %>%
  st_transform(., 27700)

#plot(stp)

endpoint<- as.data.frame(end[["osm_points"]]@coords)
class(endpoint)
edp<- endpoint%>%
  st_as_sf(., coords = c("lon", "lat"), 
           crs = 4326) %>%
  st_transform(., 27700)
#plot(edp)

```
####2.2.2 import infromation about london borough 
From [London datastore](https://data.london.gov.uk/dataset/london-borough-profiles) obtain some information about London, such as the total population density,job_density,Dwellings per hectare,Gdp and hourse price. I have Integrate them into one csv file and upload it to my [github](https://github.com/Yiran-Dou/CASA005FINAL).
```{r}
boroughdes <- read_csv("boroughdescribe.csv",
                 locale = locale(encoding = "latin1"),
                 na = "n/a")

```

###2.3 import shapefiles
####2.3.1 London borough
Got the shapefile from [London datastore](https://data.london.gov.uk/dataset/london-borough-profiles). And I have also put it in my [github](https://github.com/Yiran-Dou/CASA005FINAL).
```{r}
Londonborough <- st_read(here::here("London_Borough_Excluding_MHW.shp"))%>%
  st_transform(., 27700)
```

```{r}
plot(Londonborough)
```
combine Londonborough with the describe information:
```{r}
Londonborough <- left_join(Londonborough, boroughdes, by = c("GSS_CODE"="code"))
```

```{r}
plot(Londonborough[9])   #plot by total resident population density
```
interactivity plot
```{r}

tm_shape(Londonborough) +
  tm_polygons("total_population_density", title="population_density",palette="PuBu")

tmap_mode("view")
tmap_last()

```
####2.3.2London roads
Got roads, railways and pois from [Open street map](https://wiki.openstreetmap.org/wiki/Downloading_data).
```{r}
road <- st_read(here::here("gis_osm_roads_free_1.shp"))%>%
  st_transform(., 27700)
```

```{r}
plot(road$geometry,col=c("#99CC99"),lwd=0.5)
```
####2.3.3 London railways
```{r}
railways <- st_read(here::here("gis_osm_railways_free_1.shp"))%>%
  st_transform(., 27700)
```

```{r}
plot(railways$geometry,lwd=1.5,col=c("#FFCC99"))
```
####2.3.4 London pois
```{r}
pois <- st_read(here::here("gis_osm_pois_a_free_1.shp"))%>%
  st_transform(., 27700)
```

```{r}
plot(pois$geometry)
```
##3 Visualization
###3.1 Mark the start and end point of uber on the map
####3.1.1 start point
```{r}
tm_shape(Londonborough) +
  tm_polygons(col = NA, alpha = 0.5) +
  tm_shape(stp) +
  tm_dots(col=c("#FF9999"))
```
It is obvious that some points are beyond the boundary. Only keep the point in the range.
```{r}
stplondon <- stp[Londonborough,]

tmap_mode("view")
tm_shape(Londonborough) +
  tm_polygons(col = NA, alpha = 0.5) +
  tm_shape(stplondon) +
  tm_dots(col=c("#FF9999"))
```
####3.1.2 end point
Perform the same processing as start
```{r}
edplondon <- edp[Londonborough,]

tmap_mode("view")
tm_shape(Londonborough) +
  tm_polygons(col = NA, alpha = 0.5) +
  tm_shape(edplondon) +
  tm_dots(col=c("#339933"))
```
###3.2 Count the times in the area
####3.2.1 start time in every borough
```{r}

polygon <- Londonborough[1:33, "NAME"]
intersection <- st_intersection(x =polygon, y = stp)

# Plot intersection
plot(polygon)
plot(stp,add = TRUE)

table(intersection$NAME) 

# using dplyr
int_result_stp <- intersection %>% 
  group_by(NAME) %>% 
  count()

as.data.frame(int_result_stp)[,-3]

```
####3.2.2 end times per borough
```{r}

intersection2 <- st_intersection(x =polygon, y = edp)
plot(polygon)
plot(edp,add = TRUE)

table(intersection2$NAME) 

int_result_edp <- intersection2 %>% 
  group_by(NAME) %>% 
  count()

as.data.frame(int_result_edp)[,-3]

```
###3.3 plot by start and end counts
join data:
```{r}
Londonborough1 <- left_join(Londonborough, st_drop_geometry(int_result_stp), by = c("NAME"="NAME"))
Londonborough2 <- left_join(Londonborough1, st_drop_geometry(int_result_edp), by = c("NAME"="NAME"))
```
By start:
```{r}

tmap_mode("view")
tm_shape(Londonborough1) +
  tm_polygons("n", title="start counts",palette="OrRd")

```
By end:
```{r}


tm_shape(Londonborough2) +
  tm_polygons("n.y", title="end counts",palette="OrRd")

```
###3.4 plot the condition of borough
```{r}
tmap_mode("plot")
tm1 <- tm_shape(Londonborough) + 
  tm_polygons("job_density2018", 
              palette="OrRd")+
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("Job", position=c(0,0.85), size=1.5)

tm2 <- tm_shape(Londonborough) + 
  tm_polygons("GDP",
              palette="Reds") + 
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("GDP", position=c(0,0.85), size=1.5)

tm3 <- tm_shape(Londonborough) + 
  tm_polygons("hourse_price_mean",
              palette="PuBu") + 
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("Hourse price", position=c(0,0.85), size=1.5)

tm4 <- tm_shape(Londonborough) + 
  tm_polygons("total_population_density", 
              palette="OrRd")+
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("Job", position=c(0,0.85), size=1.5)


legend <- tm_shape(Londonborough) + 
  tm_polygons("job_density2018", 
              palette="OrRd")+
  tm_layout(legend.only = TRUE)+
  tm_shape(Londonborough) + 
  tm_polygons("GDP",
              palette="Reds")+
  tm_layout(legend.only = TRUE)

legend2<- tm_shape(Londonborough) + 
  tm_polygons("hourse_price_mean",
              palette="PuBu") + 
  tm_layout(legend.only = TRUE)+
  tm_shape(Londonborough) + 
  tm_polygons("total_population_density", 
              palette="OrRd")+
  tm_layout(legend.only = TRUE)

t<-tmap_arrange(tm1, tm2, tm3, tm4,legend, legend2, ncol=3)
t

```
start and end point counts
```{r}

tm5 <- tm_shape(Londonborough2) +
  tm_polygons("n.y",
              palette="OrRd")+
  tm_layout(frame=FALSE)+
  tm_credits("end", position=c(0,0.85), size=1.5)
tm6<- tm_shape(Londonborough1) +
  tm_polygons("n", 
              palette="OrRd")+
  tm_layout(frame=FALSE)+
  tm_credits("start", position=c(0,0.85), size=1.5)

t1<-tmap_arrange(tm5, tm6, ncol=2)
t1

```

##4.Choose spacial district to analysis
###4.1 choose city of london, westminster and barnet
as city of london, westminster and barnet show special pattern, pick it out to observe.
```{r}
Londonward <- st_read(here::here("London_Ward.shp"))%>%
  st_transform(., 27700)  
plot(Londonward)

City_of_london <- Londonward %>%
  filter(., BOROUGH=="City of London")
tmap_mode("plot")

tm_shape(City_of_london) +
  tm_polygons(col = NA, alpha = 0.5)

Barnet <- Londonward %>%
  filter(., BOROUGH=="Barnet")

tm_shape(Barnet) +
  tm_polygons(col = NA, alpha = 0.5)

Westminster <- Londonward %>%
  filter(., BOROUGH=="Westminster")
tmap_mode("plot")

tm_shape(Westminster) +
  tm_polygons(col = NA, alpha = 0.5)
```
###4.2 combine with start point
```{r}

stpcityoflondon <- stp[City_of_london,]

tmap_mode("view")
tm_shape(City_of_london) +
  tm_polygons(col = NA, alpha = 0.5) +
  tm_shape(stpcityoflondon) +
  tm_dots(col=c("#FF9999"))

stpWestminster <- stp[Westminster,]
tm_shape(Westminster) +
  tm_polygons(col = NA, alpha = 0.5) +
  tm_shape(stpWestminster) +
  tm_dots(col=c("#FF9999"))


stpbarnet <- stp[Barnet,]

tm_shape(Barnet) +
  tm_polygons(col = NA, alpha = 0.5) +
  tm_shape(stpbarnet) +
  tm_dots(col=c("#FF9999"))

```
###4.3 start and end point density
```{r}
se <- read_csv("start_end.csv",
                 locale = locale(encoding = "latin1"),
                 na = "n/a")
Londonborough3 <- left_join(Londonborough, se, by = c("area"="area"))
tmap_mode("plot")
tmap_mode("plot")
tm1 <- tm_shape(Londonborough3) + 
  tm_polygons("start", 
              palette="OrRd")+
   tm_layout(frame=FALSE)+
  tm_credits("start point density", position=c(0,0.85), size=1.5)

tm2 <- tm_shape(Londonborough3) + 
  tm_polygons("end",
              palette="Reds") + 
  tm_layout(frame=FALSE)+
  tm_credits("end point density", position=c(0,0.85), size=1.5)
t=tmap_arrange(tm1, tm2, ncol=2)
t
```

